Generalized Plackett-Luce Likelihoods

### 2023-05-09 inst/jss_4382_rebuttal.tex

1.3 Specific comments\\ S1 I'm not claiming any expertise in English
language, but occasionally, I found the manuscript difficult to
understand. For example, in the abstract it says: "The package
. . . furnishes idiom" Couldn't that simply be: "The package
. . . provides functionality"?

\textcolor{blue}{The original motivation for the package occurred when
  I had a large complicated set of Bradley-Terry observations which I
  wanted to translate into a likelihood function.  The R code became
  extremely long and unweildy, and was almost impossible to check.  It
  also became very slow to run.  What I wanted was a nice way to enter
  this kind of observations, in such a way as to maintain both
  computational efficiency [hence the use of the {\tt STL} map class]
  {\bf and} human readability.  I think that the phrase ``furnishes
  idiom" expresses this reasonably clearly.  But on reflection it does
  sound a bit odd and I have rephrased the abstract accordingly.}\\ \\

On P. 13, it says "Here we use generalized Plackett-Luce to assess the
constructors' performance". Isn't there a "model" missing?
(Plackett-Luce is just a name and cannot be used).
\textcolor{blue}{The original text was too terse and I have
  reworded.}\\ \\

S2 Figure 2: For readability, maybe use xlab = expression(p[a]) and maybe
instead of a legend, write AAB, ABA, BAA next to the line in the color
of that line.
\textcolor{blue}{Done, figure much improved}\\ \\

S3 P. 7: "https:github.com/RobinHankin/hyper2" should be
"https://github.com/RobinHankin/hyper2".
\textcolor{blue}{Corrected}\\ \\

S4 P. 7: "Wolf" should be "Wolfe".
\textcolor{blue}{Corrected.}\\ \\

S5 P. 15, maxp(const2020, n=1): What does n do?
\textcolor{blue}{Argument $n$ discussed above under Timing}\\ \\

S6 P. 16, "we see strong evidence for a real decrease in the strength": I'd tone
that down. The model is almost certainly wrong, so it does not measure
the "real decrease." Also, power is unknown, so it is difficult to say what
"evidence" there is against H0. (See, e.g., Gelman, Regression an other
stories, Chapter 4: Time to unlearn what you thought you knew about
statistics. https://statmodeling.stat.columbia.edu/2022/01/27/regression-
and-other-stories-free-pdf/)\\

\textcolor{blue}{I have not seen ROS before and have been studying
  chapter 4 as suggested.  I make some discursive responses
  here.\\ \\ I was fascinated by ``Researcher degrees of freedom''; I
  am familiar with the concept but have not seen this particular, very
  apposite, phrase used to describe it.  I would observe that my work
  is not in the same class as discussed in ``solving the Bible code
  puzzle'' [in which an unacceptably high amount of researcher DoF was
    documented].  My own work is, for example, written to
  reproducibility standards and all data and indeed code held on
  public repositories [github and CRAN], open to scrutiny.  There is
  very little ``wiggle room'' in my approach: the data is unambiguous,
  and the model (Plackett-Luce) is clear and objective.\\ \\ I would
  observe that Gelman uses the frequentist paradigm uncritically, a
  position adopted by a decreasing minority of researchers.  For
  example, the notion of ``bias'' is discussed; but many workers
  (myself included) regard the bias of an estimator as unimportant: a
  biased estimator with a small mean square error is not without
  value.\\ \\I am not sure what relevance ``power'' has here: this is
  the probability of {\em correctly} failing to reject the null when
  it is false (and here, I did not fail to reject my null: perhaps we
  should include an ensemble of Earths and visit a random one on which
  ``I'' failed to reject the null).\\ \\ The comment ``the model is
  almost certainly wrong'' is, I would argue, misleading.  All models
  are wrong.  But some are useful.  My view is that the truth of a
  statistical model [for some theological definition of truth] is a
  poor criterion: no random variable is truly Gaussian, for example;
  but the Student $t$-test is not useless as a result.  The issue is
  whether non-normality materially affects the properties of the test.
  Often this is not the case and the Gaussian approximation is
  adequate: the $t$-test is sufficiently robust to be useful in a wide
  range of situations.  True? No. Useful?  Consistent? Reliable?
  Trustworthy?  Easy to use?  Familiar?  Well-understood? Standard?
  Effective?  Powerful?  Yes.  The Student $t$-test has value, and so
  does Plackett-Luce.  My position is that the framework set out in
  the manuscript is plausible and intuitive.  I make no claim for
  absolute correctness.\\ \\ But in answer to your question, the small
  $p$-value and large likelihood ratio statistic certainly {\em is}
  evidence for a real decrease.  The hypothesis of no real strength
  change has low likelihood compared with the alternative of a drop in
  Bradley-Terry strength.  Either Plackett-Luce is seriously wrong, or
  PL is correct but some low-probability event has occurred which is
  misleading us.  Alternatively, as I suggest, Mercedes really is
  suffering from some genuine race-affecting issue [porpoising? Lewis
    Hamilton getting older?  unaccustomed F1 budget caps?].  The {\em
    strength} of this evidence is quantified by the small $p$-value
  (here less than 0.002) or alternatively a likelihood ratio of
  $e^{4.722}\simeq 112.4$, surely worthy of the epithet
  ``strong''.}\\ \\

### 2023-05-09 inst/jss_hyper3.Rnw
more work on issue #153

\begin{figure}[htbp]
  \begin{center}
<<figabbabb,echo=FALSE,fig=TRUE>>=
f_aab <- function(a){a^2 / (1 + a)}
f_aba <- function(a){a * (1 - a) / (1 + a)^2}
f_baa <- function(a){(1 - a) / (1 + a)}
p <- function(f, ...){
  a <- seq(from = 0, by = 0.005, to = 1)
  points(a, f(a) / max(f(a)), ...)
  }
plot(0:1, 0:1, xlab = expression(p[a]), ylab = "Likelihood", type = "n")
p(f_aab, type = "l", col = "black")
p(f_aba, type = "l", col = "red")
p(f_baa, type = "l", col = "blue")
text(0.8,0.8,"AAB")
text(0.8,0.5,"ABA",col="red")
text(0.8,0.15,"BAA",col="blue")
abline(h = exp(-2), lty = 2)
@
\caption{Likelihood \label{ABAetc_likelihood} functions for
observations $a\succ a\succ b$, $a\succ b\succ a$, $b\succ a\succ a$.}
\end{center}
\end{figure}

@@

The ideas presented above can easily be extended to arbitrarily large
numbers of competitors, although the analytical expressions tend to be
intractable and numerical methods must be used.  All results and
datasets presented here are maintained under version control and
available at \url{https:github.com/RobinHankin/hyper2}.  Given an
order statistic of the type considered above, the
Mann-Whitney-Wilcoxon test~\citep{mann1947,wilcoxon1945} assesses a
null of identity of underlying distributions~\citep{ahmad1996}.
Consider the chorioamnion dataset~\citep{hollander2013}, used in
\code{wilcox.test.Rd}:

>>>>

\begin{figure}[htbp]
  \begin{center}
<<figabbabb,echo=FALSE,fig=TRUE>>=
f_aab <- function(a){a^2 / (1 + a)}
f_aba <- function(a){a * (1 - a) / (1 + a)^2}
f_baa <- function(a){(1 - a) / (1 + a)}
p <- function(f, ...){
  a <- seq(from = 0, by = 0.005, to = 1)
  points(a, f(a) / max(f(a)), ...)
  }
plot(0:1, 0:1, xlab = expression(p[a]), ylab = "Likelihood", type = "n")
p(f_aab, type = "l", col = "black")
p(f_aba, type = "l", col = "red")
p(f_baa, type = "l", col = "blue")
text(0.8,0.8,"AAB")
text(0.8,0.5,"ABA",col="red")
text(0.8,0.15,"BAA",col="blue")
abline(h = exp(-2), lty = 2)
@
\caption{Likelihood \label{ABAetc_likelihood} functions for
observations $a\succ a\succ b$, $a\succ b\succ a$, $b\succ a\succ a$.  Horizontal dotted
line represents two units of support}
\end{center}
\end{figure}

@@

The ideas presented above can easily be extended to arbitrarily large
numbers of competitors, although the analytical expressions tend to be
intractable and numerical methods must be used.  All results and
datasets presented here are maintained under version control and
available at \url{https://github.com/RobinHankin/hyper2}.  Given an
order statistic of the type considered above, the
Mann-Whitney-Wilcoxon test~\citep{mann1947,wilcoxon1945} assesses a
null of identity of underlying distributions~\citep{ahmad1996}.
Consider the chorioamnion dataset~\citep{hollander2013}, used in
\code{wilcox.test.Rd}:
